{\bf Practical Concurrency Testing} \\ \normalsize \vspace{1em} \begin{tabular}{c} {\em or, How I Learned to Stop Worrying and Love the Exponential Explosion} \end{tabular}

### 2018-02-09 thesis/landslide.tex
scheudler

\subsection{Scheduler}
\subsection{Memory analysis}

@@

{\tt mem\_check\_shared\_access()} also records each such access in a per-thread-transition rbtree,
which is saved and then cleared at each preemption point.
This allows {\tt mem\_shm\_intersect()}, called at each preemption point once for each of its ancestors
($n^2$ total calls per interleaving),
to perform a set intersection to find any memory conflicts.
Any such conflicts whicih also fail a lockset (\sect{\ref{sec:landslide-lhb}})
and/or happens-before (\sect{\ref{sec:landslide-phb}}) check
are then reported as data races.
Regardless, all such conflicts are later used by DPOR (\sect{\ref{sec:landslide-dpor}}) to find dependent transition pairs.

@@

\subsubsection{Interrupts}

@@

\subsubsection{Arbiter}

@@

\subsubsection{Time Travel}

>>>>

Both simulators invoke Landslide once per instruction and once per memory read or write.
% just before..
The entry point is the aptly-named {\tt landslide\_entrypoint()} in {\tt landslide.c},
which then dispatches to various other modules their respective analyses, described as follows.

\subsection{Execution Tree}
\label{sec:landslide-save}

\subsection{Scheduler}

The Landslide scheduler, which lives in {\tt schedule.c}, has two main duties:
to maintain an accurate representation of all the existing threads on the simulated system
and track what concurrency-relevant actions each is performing at any given time,
and to orchestrate the sequence of timer interrupts necessary
to cause the simulated system to context switch to any given thread at any given time.
System-wide state is stored in a single {\tt struct sched\_state},
including the thread queues (runqueue, deschedule queue, and sleep queue),
while per-thread state is stored in {\tt struct agent}s (named after the terminology of \cite{dbug-ssv}) which live on said queues.

It has one main entrypoint, {\tt sched\_update()}, in which both the state machine is updated and scheduling decisions are made.
The interface also offers helper functions for finding and manipulating {\tt agent}s,
and {\tt sched\_recover()}, which prepares the scheduler to force a new thread to run
after a time travel (\sect{\ref{sec:landslide-timetravel}}).

\subsubsection{State machine}

The first part of {\tt sched\_update()} is to update the state machine of thread actions and runnability.
Much of this functionality is found in
{\tt sched\_update\_kern\_state\_machine()}
and
{\tt sched\_update\_user\_state\_machine()}.
The current intruction pointer is compared against the known locations of
the mutex API, system calls, runnable/descheduling {\tt tell\_\allowbreak{}landslide} annotations, and so on,
and locksets, action flags, and runqueue membership are updated accordingly.

\subsubsection{Interrupt injection}

The second part of {\tt sched\_update()},
conditional on the arbiter identifying preemption points (\sect{\ref{sec:landslide-arbiter}}),
manages timer interrupts to switch to a desired thread.
Whenever a preemption point is reached,
the scheduler first creates a checkpoint in the execution tree (\sect{\ref{sec:landslide-save}}),
asks the arbiter which thread to run next,
and if that thread is different from the current one,
forces the kernel into its timer interrupt handler (\sect{\ref{sec:landslide-interrupce}}).

Because the kernel is part of the system being tested, Landslide can't necessarily always switch directly to a specific thread,
but rather must keep triggering context switches until the desired thread is reached;
any mechanism to tell the kernel which thread we want would necessarily involve modifying the code being testsed
and hence possibly obscuring bugs or introducing new ones.%
\footnote{For userspace testing, where I supply a pre-annotated reference kernel,
such an approach would be more straightforward,
but the kernel-testing repeated-context-switch approach infrastructure was already in place
and it was easier to reuse that than to add more code.}

The scheduler marks up to one thread as the ``schedule target'',
which when set makes Landslide wait until that thread is reached before looking for more preemption points,
so the kernel may finish its context switches undisturbed.
Whenever the schedule target is set and the end of the context switcher is reached,
if the schedule target is not the current thread,
the scheduler repeats this process until it is.%
\footnote{Note that this ``loop'' is not structured as an explicit loop in Landslide's code,
but rather as part of the state machine which updates each time a new instruction is traced.}

\subsection{Memory analysis}

@@

{\tt mem\_check\_shared\_access()} also records each such access in a per-thread-transition rbtree,
which is saved and then cleared at each preemption point.
This allows {\tt mem\_shm\_\allowbreak{}intersect()}, called at each preemption point once for each of its ancestors
($n^2$ total calls per interleaving),
to perform a set intersection to find any memory conflicts.
Any such conflicts which also fail a lockset
and/or happens-before check (\sect{\ref{sec:landslide-datarace}})
are then reported as data races.
Regardless, all such conflicts are later used by DPOR (\sect{\ref{sec:landslide-dpor}}) to find dependent transition pairs.

@@

\subsubsection{Interrupts}
\label{sec:landslide-interrupce}

@@

\subsubsection{Arbiter}
\label{sec:landslide-arbiter}

@@

\subsubsection{Time Travel}
\label{sec:landslide-timetravel}
